---
title: "HW13: LM and LMM"
output: pdf_document
date: "2024-11-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: Chimp Brain Polynomial

### a

```{r}
#read in and standard plot
chimpcortex <- read.csv(file = "chimpcortex.csv")
plot(chimpcortex)

```

### b

```{r}
# Extract the response variable y and predictor variable age
y <- chimpcortex$synapses
age <- chimpcortex$age

# Construct the design matrix with polynomial terms
X <- as.matrix(cbind(1, age, age^2, age^3))

# Calculating beta using the normal equation
beta.hat <- solve(t(X) %*% X) %*% t(X) %*% y

# Reporting the estimated beta values
beta.hat

# Plot the data
plot(chimpcortex$age, chimpcortex$synapses, 
     xlab = "Age (years)", 
     ylab = "Synapses per 100 cubic micrometers", 
     main = "Synaptic proliferation in chimpanzee motor cortex")

# Add the fitted curve
curve(beta.hat[1] + beta.hat[2]*x + beta.hat[3]*x^2 + beta.hat[4]*x^3, 
      add = TRUE, col = "red", lwd = 2)


# Fit the model using lm()
model <- lm(synapses ~ age + I(age^2) + I(age^3), data = chimpcortex)

# Reporting the estimates
summary(model)$coefficients

```

## Question 2: Mouse Body Weight

### a

```{r}
# Load the data
mousebw <- read.csv(file = "mousebw.csv")

# Fit the linear model
lm_model <- lm(bw ~ diet, data = mousebw)

# Report the point estimate for the diet effect
diet_effect <- summary(lm_model)$coefficients["diet", "Estimate"]
round(diet_effect, 2) #2 for the decimal point


```

### b

```{r}
# Load necessary library
library(lme4)

# Fit the linear mixed model
fitr <- lmer(bw ~ diet + (1 | strain), data = mousebw)
```

 Write out the regression model

$\text{bw}\_i = \beta\_0 + \beta\_1 \cdot \text{diet}\_i + \epsilon\_i$

Where:

$(\text{bw}_i)$ is the body weight of the (i)-th mouse.

$(\beta_0)$ is the intercept, representing the average body weight of mice on the reference diet (e.g., normal diet).

$(\beta_1)$ is the coefficient for the diet effect, representing the change in body weight associated with the high-fat diet compared to the normal diet.

$(\text{diet}_i)$ is a binary variable indicating the diet of the (i)-th mouse (e.g., 0 for normal diet, 1 for high-fat diet).

$(\epsilon_i)$ is the error term, capturing the variation in body weight not explained by the diet.

```{r}
# Call summary() on the fitted model and report the point estimate for the diet effect
summary(fitr)$coefficients["diet", "Estimate"]

# Calculate the (ICC)
var_strain <- as.numeric(VarCorr(fitr)$strain[1])
var_residual <- attr(VarCorr(fitr), "sc")^2
ICC <- var_strain / (var_strain + var_residual)
ICC

# Get the shrinkage estimates for the strain effects
shrinkage_estimates <- ranef(fitr)$strain
shrinkage_estimates
```

Linear Model: This part fits a simple linear model to estimate the effect of diet on body weight.

Linear Mixed Model: This part accounts for the additional dependency structure due to different strains by including a random intercept for strain.
