---
title: "HW9"
format: pdf
editor: visual
---

## Homework 9 - Multiple Testing

### Question 1: Differential Gene Expression

a.i

```{r}
#read in the HW file
pmix <- scan(file = "pmix.txt")

#histogram of data displayed as proportions
pmix.hist <- hist(pmix)
pmix.hist$density <- pmix.hist$counts / sum(pmix.hist$counts)
plot(pmix.hist, freq = F)

```

There is a peak on the far left side, so this is a sign there is signal in the data. This resembles more of a exponential distribution

a.  ii

```{r}
#generate even points from 0 to 1
x <- ppoints(1000)
x <- sort(x)

#sort the txt file points
y <- pmix[order(pmix)]

#plotting the null points (900)
plot(x[1:900], y[1:900], main = "Q-Q Plot for Null",
     xlab = "Theoretical Quantiles", ylab = "Observed P-values", pch = 19, cex = 0.5)+
abline(0, 1, col = "blue", lty = 2)

# Plot for the whole set of 1,000 p-values
plot(x, y, col=ifelse(x<=0.05, "red", "black"), main = "Q-Q Plot for All 1k P-values",
     xlab = "Theoretical Quantiles", ylab = "Observed P-values", pch = 19, cex = 0.5)+
abline(0, 1, col = "blue", lty = 2)

```

a.  iii these plots look the same so something is wrong i think

```{r}
# seperate and sort the null points 
non.null <- pmix[1:100]
null.p <- pmix[101:1000]

non.null <- sort(non.null)
null.p <- sort(null.p)

trans.non.null <- -log10(non.null)
trans.null.p <- -log10(null.p)

ppx <- ppoints(900)
ppx <- -log10(ppx)

# plot the null with the transformed points

plot(trans.null.p, ppx, main = "Q-Q Plot for Null",
     xlab = "-log10 Theoretical Quantiles", ylab = "-log10 Observed P-values", pch = 19, cex = 0.5, xlim = c(0,6), ylim = c(0,6))+
abline(0, 1, col = "blue", lty = 2)
  
# log10 transformation
x_transformed <- -log(x)
y_transformed <- -log(y)

# plot the null with the transformed points

#  plot all 1k points with transformed points 
plot(x_transformed, y_transformed, col=ifelse(x<=0.05, "red", "black"), 
     main = "Q-Q Plot for All 1k P-values",
     xlab = "-log10(Theoretical Quantiles)", ylab = "-log10(Observed P-values)", 
     pch = 19, cex = 0.5)+
abline(0, 1, col = "blue", lty = 2)



```

a.  iv The researchers transformed the data for better visualization of the significant points and their proportion to the null values. This plot also improves on seeing how far from the null the points are, not just the significance level.

### Question 2 MCC Math

```{r}

TP <- sum(pmix <= 0.05)

FP <- 45

FN <- 45

TN <- sum(pmix>=0.05)

# MCC Math

top <- (TP*TN)-(FP*FN)

srt <- (TP+FP)*(TP+FN)*(TN+FP)*(TN+FN)

bottom <- sqrt(srt)

top/bottom

```

### Question 2:C ii

```{r}
# BF correction 
bon.f.pvals <- p.adjust(pmix, method = "bonferroni")

corrected.TP <- sum(bon.f.pvals <=0.05)
corrected.TN <- sum(bon.f.pvals >=0.05)
corrected.FP <- 100 - corrected.TP 
corrected.FN <- corrected.TN - 900

cat("The corrected values are:", "TP =", corrected.TP, ", TN =", corrected.TN, ", FP =", corrected.FP, "FN = ", corrected.FP) 

```

### Question 2D  Benjamini-Hochberg False Discovery Rate (BH-FDR)

```{r}

fdr.p.vals <- p.adjust(pmix, method = "fdr")

sig.fdr.tp <- fdr.p.vals <= 0.05
sig.fdr.tn <- fdr.p.vals >= 0.05
sig.fdr.fn <- (sum(sig.fdr.tn)/length(fdr))


```





